---
title: "RWorksheet_Octavio#4b"
output: pdf_document
date: "2023-11-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#1
    vectorA <-c (1,2,3,4,5)
    seroMatrix<- matrix(0,nrow=5, ncol=5)
   
    for (i in 1:5){
    for (j in 1:5){
   
    hat <- abs(vectorA[i]-j)
   
    cat(hat," ")
 }
    cat ("\n")
}    

#2
    for (i in 1:5){
    hat <- rep ("*",i)
    print(hat)
}
     
#3 Get an input from the user to print the Fibonacci sequence starting from the 1st input up to 500. Use repeat and break statements. Write the R Scripts and its output.
```{r} 
   userInput <- as.integer(readline("Enter starting number for Fibonacci sequence: "))

a <- userInput
b <- 0


cat("Fibonacci sequence starting from", userInput, ":\n")

repeat {
  
  next_num <- a + b
  
  if (next_num > 500){
    break
  }
  cat(next_num, " ")
  a <- y
  b <- next_num
}  
```
#4 Import the dataset as shown in Figure 1 you have created previously.

#4a What is the R script for importing an excel or a csv file? Display the first 6 rows of the dataset? Show your codes and its result
```{r}
prevData <- read.csv("householdData.csv")
head(prevData)
```

#4b Create a subset for gender(female and male). How many observations are there in Male? How about in Female? Write the R scripts and its output.
```{r}
males <- prevData[prevData$Gender == "M",]
males

females <- prevData[prevData$Gender == "F",]
females

numofMale <- nrow(males)
numofMale

numofFem <- nrow(females)
numofFem
```






#5 The monthly income of Dela Cruz family was spent on the following:

#a Create a piechart that will include labels in percentage.Add some colors and title of the chart. Write the R scripts and show its output.

```{r}
data <- c(60, 10, 5, 25)
labels <- c("Food", "Electricity", "Savings", "Miscellaneous")
colors <- c("orange", "skyblue", "violet", "red")
pie(data, labels = labels, col = colors, main = "My Pie Chart")
```
#6 Use the iris dataset.

#a Check for the structure of the dataset using the str() function. Describe what you have seen in the output.

```{r}
data("iris")

species_count <- table(iris$Species)
colors <-c("yellow", "green", "pink")
pie(species_count, labels = levels(iris$Species), col = colors, main = "Iris Species Distribution")
```    
#b. Create an R object that will contain the mean of the sepal.length,sepal.width,petal.length,and petal.width. What is the R script and its result?
```{r}






```
#c. Create a pie chart for the Species distribution. Add title, legends, and colors. Write the R script and its result.
```{r}
data("iris")
species_count <- table(iris$Species)
colors <- c("violet", "orange", "maroon")
pie(species_count, labels = levels(iris$Species), col = colors, main = "Iris Species Distribution")
legend("topright", legend = levels(iris$Species), fill = colors, title = "Species")
```
#d. Subset the species into setosa, versicolor, and virginica. Write the R scripts and show the last six (6) rows of each species.
```{r}
data(iris)
setosa_data <- subset(iris, Species == "setosa")
versicolor_data <- subset(iris,)